---
title: "EDA House Prices"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Summary

This is an EDA for the **kaggle** competition [House Prices: Advanced Regression Techniques](https://www.kaggle.com/c/house-prices-advanced-regression-techniques>). 



```{r}
#install.packages("moments")

library(ggplot2) # Data visualization
library(readr) # CSV file I/O, e.g. the read_csv function
library(moments)
library(gridExtra)
library(caret)
library(corrplot)
```

```{r}
set.seed(1234)
```

### Loading and preprocessing the data

You can also embed plots, for example:

```{r}
train <- read.csv("../input/train.csv")
```

These are detected as numeric but they are categorical variables:

```{r}
train$MSSubClass <- as.factor(train$MSSubClass)
train$OverallQual <- as.factor(train$OverallQual)
train$OverallCond <- as.factor(train$OverallCond)
#train$YearBuilt <- as.factor(train$YearBuilt)
#train$YearRemodAdd <- as.factor(train$YearRemodAdd)
```

```{r}
numeric_var <- names(train)[which(sapply(train, is.numeric))]
factor_var <- names(train)[!(names(train) %in% numeric_var)]
```


### First look to the data and getting some summary information

```{r}
str(train)
```

There are `r nrow(train)` data rows and `r ncol(train)` variables 

```{r}
summary(train)
```

### Check NA values
we can see some NA values in variables: 

```{r}
which(apply(train, 2, function(x) any(is.na(x)))==TRUE)
all_vars_napercent <- apply(train, 2, function(x) round(sum(is.na(x))/nrow(train), 3))
naVars <- names(train)[which(apply(train, 2, function(x) round(sum(is.na(x))/nrow(train), 3))>0)]
sort(all_vars_napercent[naVars], decreasing = TRUE)
```

Some variables like *PoolQC* , *MiscFeature*, *Alley*, *Fence* has a very hight ratio of NA values. 

We can see that all of the `r nrow(train)` row have one ore more NA values:

```{r}
sum(apply(train, 1 , function(x) sum(is.na(x))) == 0)
```

Alternately 

```{r}
sum(complete.cases(train))
```

Although only `r mean(is.na(train))*100`% of the values of the data are missing values 

We can check this and get more detailed information using specialized imputation packages like [MICE](https://www.analyticsvidhya.com/blog/2016/03/tutorial-powerful-packages-imputing-missing-values/)

```{r}
library(VIM)
mice_plot <- aggr(train, col=c('navyblue','yellow'),
                    numbers=TRUE, sortVars=TRUE,
                    labels=names(train), cex.axis=.7,
                    gap=3, ylab=c("Missing data","Pattern"))
```

If we want to use some machine learning models like linear regression we will need to impute values for these NA

### Check variables with zero o near zero variance 

Let's see what variables have small or zero variance being very conservative with this:
```{r}
vars_variance <- nearZeroVar(train, saveMetrics = TRUE, freqCut = 99/1, uniqueCut = 1)
vars_variance[vars_variance[,"zeroVar"] + vars_variance[,"nzv"] > 0, ] 
```

There aren't any variable with zero variance. We will have to examine in detail these variables near zero variance to see if we can eliminate some of them.

### Check outcome variable

The outcome variable to predict is SalePrice. We will take a look to the values of this variable in detail for the training data:

```{r}
summary(train$SalePrice)
boxplot(train$SalePrice)
```

We can see SalePrice is right skewed and don't follow a normal distribution

```{r}
skewness(train$SalePrice)
shapiro.test(train$SalePrice)
qqnorm(train$SalePrice)
qqline(train$SalePrice, col = 2)
```


We can try to check what happens with a log transformation of SalePrice:
```{r}
logSalePrice <- log(train$SalePrice)
hist(logSalePrice)
skewness(logSalePrice)
shapiro.test(logSalePrice)
qqnorm(logSalePrice)
qqline(logSalePrice, col = 2)
```

It looks like it follows a more normal distribution but it has some right skew too, and its not completely normal as shapiro test show: if we consider a 0.05 level of significance we must reject the null hipotesys which consideres the data follows a normal distribution.

Lets take a look to categorical variables
```{r}
plotHist <- function(data_in, i) {
  data <- data.frame(x=data_in[[i]])
  p <- ggplot(data=data, aes(x=factor(x))) + stat_count() + xlab(colnames(data_in)[i]) + theme_light() + 
    theme(axis.text.x = element_text(angle = 90, hjust =1))
  return (p)
}

doPlots <- function(data_in, fun, ii, ncol=3) {
  pp <- list()
  for (i in ii) {
    p <- fun(data_in=data_in, i=i)
    pp <- c(pp, list(p))
  }
  do.call("grid.arrange", c(pp, ncol=ncol))
}


plotDen <- function(data_in, i){
  data <- data.frame(x=data_in[[i]], SalePrice = data_in$SalePrice)
  p <- ggplot(data= data) + geom_line(aes(x = x), stat = 'density', size = 1,alpha = 1.0) +
    xlab(paste0((colnames(data_in)[i]), '\n', 'Skewness: ',round(skewness(data_in[[i]], na.rm = TRUE), 2))) + theme_light() 
  return(p)
   
}


```

```{r}
doPlots(train[,factor_var], fun = plotHist, ii = 1:4, ncol = 2)
```

```{r}
doPlots(train[,factor_var], fun = plotHist, ii  = 5:9, ncol = 2)
```

```{r}
doPlots(train[,factor_var], fun = plotHist, ii  = 10:14, ncol = 2)
```

```{r}
doPlots(train[,factor_var], fun = plotHist, ii  = 15:20, ncol = 2)
```

```{r}
doPlots(train[,factor_var], fun = plotHist, ii  = 21:28, ncol = 2)
```

```{r}
doPlots(train[,factor_var], fun = plotHist, ii  = 29:38, ncol = 2)
```

```{r}
doPlots(train[,factor_var], fun = plotHist, ii  = 39:46, ncol = 2)
```

Mostramos las gráficas de densidad para las variables continuas

```{r}
doPlots(train[,numeric_var], fun = plotDen, ii = 2:9, ncol = 2)
```

```{r}
doPlots(train[,numeric_var], fun = plotDen, ii = 10:17, ncol = 2)
```

```{r}
doPlots(train[,numeric_var], fun = plotDen, ii = 18:25, ncol = 2)
```

```{r}
doPlots(train[,numeric_var], fun = plotDen, ii = 26:35, ncol = 2)
```

### First Data Cleaning

### Identify correlations

We take a look to the correlation of the numerical variables (we omit the Id variable too):

```{r}
cont_vars <- numeric_var[2:length(numeric_var)]
correlations <- cor(na.omit(train[, cont_vars]))
corrplot(correlations, method="square")
```

Let's see what variables are more correlated:

```{r}
high_correlations_index <- apply(correlations, 1, function(x) sum( abs(x) > 0.3 ) > 1) 
high_correlations <- correlations[high_correlations_index, high_correlations_index]
corrplot(high_correlations, method="square")
```

The hight correlated vars are these:

```{r}
cont_vars[high_correlations_index]
```


### References
+ [Tutorial on 5 Powerful R Packages used for imputing missing values](https://www.analyticsvidhya.com/blog/2016/03/tutorial-powerful-packages-imputing-missing-values/)
+ [Missing Value Treatment](http://r-statistics.co/Missing-Value-Treatment-With-R.html)
+ [Interesting graph functions](https://www.kaggle.com/notaapple/house-prices-advanced-regression-techniques/detailed-exploratory-data-analysis-using-r/notebook)
+ [Better Understand Your Data in R Using Visualization](http://machinelearningmastery.com/data-visualization-in-r/)
